Inferring Graphics Programs from Images

### 2017-05-15 paper.tex
similarity in program space figure
responses to comments

\section{Neural architecture for inferring drawing execution traces}

@@

\remark{Planning to move the description of SMC / beam search up here, too?}

@@

This network suffices to ``derender'' images like those shown in
Figure~\ref{trainingData}.  We can perform a beam search decoding to
recover what the network thinks is the most likely execution trace for
images like these. But, if the network makes a mistake (predicts an
incorrect line of code), it has no way of recovering from the error.
In order to derender an image with $n$ objects, it must correctly
predict $n$ drawing commands -- so its probability of success will
decrease exponentially in $n$, assuming it has any nonzero chance of
making a mistake.  For added robustness as $n$ becomes large, we treat
the neural network outputs as proposals for a SMC sampling scheme.
For the SMC sampler, we use pixel wise distance as a surrogate for a
likelihood function; see supplement. Figure~\ref{syntheticResults}
compares the neural network with SMC against the neural network by
itself or SMC by itself.  Only the combination of the two passes a
critical test of generalization: when trained on images with $\leq 8$
objects, it successfully parses scenes with many more objects than the
training data.
\begin{wrapfigure}{R}{0.5\textwidth}
  \begin{center}
    \begin{minipage}[t]{2.2cm}\fbox{\includegraphics[width=2cm]{figures/randomScene-58-7.png}}\end{minipage}
    \begin{minipage}[t]{2.2cm}\fbox{\includegraphics[width=2cm]{figures/randomScene-33-5.png}}\end{minipage}
    \begin{minipage}[t]{2.2cm}\fbox{\includegraphics[width=2cm]{figures/randomScene-25-4.png}}\end{minipage}
%    \begin{minipage}[t]{2cm}\includegraphics[width=2cm]{figures/randomScene-26-7.png}\end{minipage}
%    \begin{minipage}[t]{2cm}\includegraphics[width=2cm]{figures/randomScene-19-3.png}\end{minipage}
  \end{center}
  \caption{Network is trained to infer execution traces for figures like the three shown above.}\label{trainingData}
\end{wrapfigure}

@@

Given the DSL and a trace $T$, we want a program that evaluates to $T$
and also minimizes some measure of program cost:
\begin{equation}
  \text{program}(T) = \argmin_{\substack{p\in \text{DSL}\\p \text{ evaluates to } T}} \text{cost}(p)\label{programObjective}
\end{equation}
An intuitive measure of program cost is its length.  We define the
cost of a program to be the number of statements it contains, where a
statement is a ``Command'' in Table~\ref{DSL}.
\remark{The flow here is a bit off/backwards. ``We want a program that evaluates to $T$ and also minimizes some measure of program cost''---why do we care about cost? It'd be better to start by making a ``Bayesian Occam's razor'' appeal (e.g. the most compact/general program is the more likely explanation) and then saying that one way to do this is to minimize a cost function which is proportional to program length.}

The constrained optimization problem in
equation~\ref{programObjective} is intractable in general, but there
exist efficient-in-practice tools for finding exact solutions to
program synthesis problems like these. We use the state-of-the-art Sketch
tool~\cite{solar2008program}. Describing Sketch's program synthesis
algorithm is beyond the scope of this paper; see supplement.  At a
high level, Sketch takes as input a space of programs, along with a
specification of the program 's behavior and optionally a cost
function.  It translates the synthesis problem into a constraint
satisfaction problem, and then uses a quasiboolean solver to find a
minimum cost program satisfying the specification.  In exchange for
not having any guarantees on how long it will take to find a minimum
cost solution, it comes with the guarantee that it will always find a
globally optimal program.

Why synthesize a graphics program,
if the execution trace already suffices to recover the objects in an image?
Within our domain of hand-drawn figures, graphics program synthesis has several important uses:
\subsection{Extrapolating figures}
Having access to the source code of a graphics program facilitates coherent, high-level edits to the figure generated by that program. 
For example,
we can change all of the circles to squares,
were make all of the lines be dashed.
We can also \textbf{extrapolate} figures
by increasing the number of times that loops are executed.
\input{extrapolations.tex}

@@

\subsection{Correcting errors made by the neural network}

@@

Let $L(\cdot | \cdot):\text{image}^2\to \mathcal{R}$ be our likelihood
function: it takes two images, an observed target image and a
hypothesized program output, and gives the likelihood of the observed
image conditioned on the program output.
Write $\hat{T}(I)$ for the trace the model predicts for image $I$.

We can extract a few basic features of a program, like its size or how many
loops it has,
and use these features to help predict
whether a trace is the correct explanation for an image.
\begin{equation}
\hat{T}(I) = \argmax_{T} L(I | \text{render}(T)) + \theta\cdot  \phi \left( \text{program}(T) \right)
\end{equation}
where $\phi (\cdot)$ is a feature extractor for programs.
This is equivalent to doing MAP inference in a generative model where the program
is first drawn from a log linear distribution $\propto \exp (\theta\cdot \phi (\text{program}))$,
than the program is executed deterministically,
and then we observe a noisy version of the program's output,
where $L$ is the noise model.

@@

\pagebreak
  \section{Preliminary Synthesis results} 
\input{synthesizerOutputs.tex}

>>>>

\section{Neural architecture for inferring drawing execution traces}\label{neuralNetworkSection}

@@

@@

This network suffices to ``derender'' images like those shown in
Figure~\ref{trainingData}.  We can perform a beam search decoding to
recover what the network thinks is the most likely execution trace for
images like these, recovering traces maximizing $\probability_\theta
[T|I]$. But, if the network makes a mistake (predicts an incorrect
line of code), it has no way of recovering from the error.  In order
to derender an image with $n$ objects, it must correctly predict $n$
drawing commands -- so its probability of success will decrease
exponentially in $n$, assuming it has any nonzero chance of making a
mistake.  For added robustness as $n$ becomes large, we treat the
neural network outputs as proposals for a SMC sampling scheme.  For
the SMC sampler, we use pixel wise distance as a surrogate for a
likelihood function. The SMC sampler is designed to produce samples
from the distribution $\propto L(I|\text{render}(T))
\probability_\theta[T|I]$, where $L(\cdot | \cdot):\text{image}^2\to
\mathcal{R}$ uses the distance between two images as a proxy for a
likelihood.

Figure~\ref{syntheticResults}
compares the neural network with SMC against the neural network by
itself or SMC by itself.  Only the combination of the two passes a
critical test of generalization: when trained on images with $\leq 8$
objects, it successfully parses scenes with many more objects than the
training data.
\begin{wrapfigure}{R}{0.5\textwidth}
  \begin{center}
    \begin{minipage}[t]{2.2cm}\fbox{\includegraphics[width=2cm]{figures/randomScene-58-7.png}}\end{minipage}
    \begin{minipage}[t]{2.2cm}\fbox{\includegraphics[width=2cm]{figures/randomScene-33-5.png}}\end{minipage}
    \begin{minipage}[t]{2.2cm}\fbox{\includegraphics[width=2cm]{figures/randomScene-25-4.png}}\end{minipage}
%    \begin{minipage}[t]{2cm}\includegraphics[width=2cm]{figures/randomScene-26-7.png}\end{minipage}
%    \begin{minipage}[t]{2cm}\includegraphics[width=2cm]{figures/randomScene-19-3.png}\end{minipage}
  \end{center}
  \caption{Network is trained to infer execution traces for figures like the three shown above.}\label{trainingData}
\end{wrapfigure}

@@

Given the DSL and a trace $T$, we want to recover a program that both evaluates to $T$
and, at the same time, is the ``best'' explanation of $T$.
For example, we might prefer more general programs or, in the spirit of Occam's razor,
prefer shorter programs.
We wrap these intuitions up into a cost function over programs,
and seek the minimum cost program consistent with $T$:
\begin{equation}
  \text{program}(T) = \argmin_{\substack{p\in \text{DSL}\\p \text{ evaluates to } T}} \text{cost}(p)\label{programObjective}
\end{equation}
We define the
cost of a program to be the number of statements it contains, where a
statement is a ``Command'' in Table~\ref{DSL}.
%\remark{The flow here is a bit off/backwards. ``We want a program that evaluates to $T$ and also minimizes some measure of program cost''---why do we care about cost? It'd be better to start by making a ``Bayesian Occam's razor'' appeal (e.g. the most compact/general program is the more likely explanation) and then saying that one way to do this is to minimize a cost function which is proportional to program length.}

The constrained optimization problem in
equation~\ref{programObjective} is intractable in general, but there
exist efficient-in-practice tools for finding exact solutions to
program synthesis problems like these. We use the state-of-the-art Sketch
tool~\cite{solar2008program}. Describing Sketch's program synthesis
algorithm is beyond the scope of this paper; see supplement.  At a
high level, Sketch takes as input a space of programs, along with a
specification of the program 's behavior and optionally a cost
function.  It translates the synthesis problem into a constraint
satisfaction problem, and then uses a SAT solver to find a
minimum cost program satisfying the specification.  In exchange for
not having any guarantees on how long it will take to find a minimum
cost solution, it comes with the guarantee that it will always find a
globally optimal program.

Why synthesize a graphics program,
if the execution trace already suffices to recover the objects in an image?
Within our domain of hand-drawn figures, graphics program synthesis has several uses:
\subsection{Extrapolating figures}
Having access to the source code of a graphics program facilitates coherent, high-level edits to the figure generated by that program. 
For example,
we can change all of the circles to squares,
were make all of the lines be dashed.
We can also \textbf{extrapolate} figures
by increasing the number of times that loops are executed.
%\input{extrapolations.tex}

@@

\includegraphics[width = 8in]{figures/similarity.png}

\subsection{Correcting errors made by the neural network}
The program synthesizer can help correct errors from  the neural network by favoring execution traces which lead to more concise or general programs.
For example, one generally prefers figures with neatly aligned objects over figures whose parts are slightly misaligned -- and nice alignment lends itself to short programs.
Similarly, figures often have repeated parts,
which the program synthesizer might model as a reflection.
So, if one is considering several candidate traces proposed by the neural network,
we might prefer traces whose best programs have desirable features like being short or having iterated structures.

Concretely we implemented the following scheme: the neurally guided sampling scheme of section~\ref{neuralNetworkSection} for image $I$ produces (samples) candidate traces, $\mathcal{F}(I)$.
Instead of predicting the most likely trace in $\mathcal{F}(I)$ according to the neural network,
we can take into account the programs that best explain the traces. 
Writing $\hat{T}(I)$ for the trace the model predicts for image $I$,
\begin{equation}
\hat{T}(I) = \argmax_{T\in \mathcal{F}(I)} L(I | \text{render}(T)) \times\probability_{\theta_p} [ \text{program}(T)] 
\end{equation}
where $\probability_{\theta_p} [\cdot]$ is a prior probability
distribution over programs parameterized by $\theta_p$.
This is equivalent to doing
MAP inference in a generative model where the program is first drawn
from $\probability_{\theta_p} [\cdot]$, then the program is executed deterministically,
and then we observe a noisy version of the program's output, where $L$
is the noise model.

Given a corpus of graphics program synthesis problems with annotated ground truth traces (so pairs of $(I,T)$),
we find a maximum likelihood estimate of $\theta_p$:
\begin{equation}
  \theta_p^* = \argmax_{\theta_p} \expect \left[ \log \frac{\probability_{\theta_p} [\text{program}(T)]\times L(I|\text{render}(T))}{\sum_{T'\in \mathcal{F}(I)} \probability_{\theta_p} [\text{program}(T')]\times L(I|\text{render}(T'))} \right]
\end{equation}
where the expectation is taken both over the model predictions and the
$(I,T)$ pairs in the training corpus.  We define $\probability_{\theta_p}
[\cdot]$ to be a log linear distribution $\propto \exp
(\theta\cdot \phi(\text{program}))$, where $\phi(\cdot)$ is a feature
extractor for programs.  We can extract a few basic features of a
program, like its size or how many loops it has, and use these
features to help predict whether a trace is the correct explanation
for an image.

@@

@@

\pagebreak
  \section{Preliminary Synthesis results} 
%\input{synthesizerOutputs.tex}
